PhD Research Proposal: Molecular Level Design of Nanoporous Lyotropic Liquid Crystal Membranes for Aqueous Separations \\ \vspace{0.5cm} \large Advisors: Michael Shirts and Richard Noble

### 2019-04-19 Ben_Manuscripts/comps/Outline.tex
UPdates to comps outline

  % re-word. This paragraph is verbatim from structure paper
  More highly selective nanoporous membranes would be extremely useful
  in performing complex aqueous separations with seawater and various
  types of wastewater.
  \begin{itemize}
    \item For example, Sodium chloride and boron in seawater 
    \cite{fritzmann_state---art_2007} and organic micropollutants found in
    municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006}
    represent just a few of the diverse contaminants of water sources. 
    \item By efficiently separating contaminants from feed solutions with
    highly selective membranes, it is possible to reduce the number of 
    required membrane passes and post-treatment steps needed for a given 
    filtration process \cite{werber_materials_2016}, thus lowering energy
    requirements. 
    \item Additionally, one can also extract valuable resources from the 
    feed streams. For example, flowback water produced during hydraulic
    fracturing of shale formations contains dissolved species such as acetate
    whose extraction has economic value \cite{dischinger_application_2017}.
  \end{itemize}

  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing membrane
  filtration processes that can be used to separate solutes on the order of
  1 nm in size and smaller, including ions.
  \begin{itemize}  
    \item Both apply hydraulic pressure to the feed solution in order to 
    overcome osmotic pressure and force water and unfiltered components 
    through the membrane.
    \item RO membranes are typically thin film composite with a porous mechanical 
    support layer and a thin but dense polymer matrix active layer where separations
    occur.\cite{jeong_interfacial_2007}
    \item RO separates solutes based on the solute's ability to dissolve into
    and diffuse through the tortuous pathways available in dense active layer.
    \item RO offers high selectivity at the cost of relatively high energy 
    requirements since one must apply a large hydraulic pressure in order to 
    achieve an economical flux.  % get number
    \item In contrast to RO membranes, NF membranes have explicit pores on the
    order of 1 nm in size. 
    \item Typically, separations are achieved based on size exclusion and
    charge exclusion if the membrane surface is charged.
    \item NF membranes require significantly less applied pressure in order
    to achieve solute flux comparable to RO.
    \item Unfortunately, conventional synthesis processes, such as 
    phase-inversion\cite{smolders_microstructures_1992} are stochastic in
    nature which yields pores that are polydisperse in size.\cite{werber_materials_2016} % name some synthesis processes
    \item Pore size polydispersity is detrimental to membrane selectivity
  \end{itemize}

@@

  Nanostructured membranes attempt to overcome the permeability-selectivity 
  tradeoff through molecular level design.
  \begin{itemize}
	\item Graphene sheets, carbon nanotubes (CNTs) and zeolites are three
	highly studied nanostructured technologies.
  \end{itemize}

  Graphene membranes are an extremely active are of research because 
  they offer potential for extremely high permeability membranes.
  \begin{itemize}
    \item 2D materials
    \item Functionalization?
    \item Easy to introduce defects
    \item Multi-layered
  \end{itemize}

  Carbon nanotubes have shown promise due to unprecedentedly fast
  water transport. 
  \begin{itemize}
    \item Functionalization of nanotubes
    \item Aligned nanotubes difficult to synthesize
  \end{itemize}

  Zeolite-coated ceramic membranes offer the potential for permeabilities
  comparable to ultrafiltration with selectivities as good as NF and RO. 
  \begin{itemize}
    \item Zeolites have highly uniform nm-sized crystalline structures with
    cage-like cavities that allow movement and trapping of small solutes. \cite{pendergast_review_2011}
    \item The crystalline frameworks are typically formed by networks of silicon
    and aluminum each attached to 4 oxygen atoms in a tetrahedral arrangement. 
    \item One can replace the silicon and aluminum atoms via ion exchange 
    in order to control the size of the cavities and hence its molecular-seiving properties.
    \item A number of studies have tested the permeability and sodium salt
    rejection of various zeolite membranes, however none have fully
    overcome the permeability-selectivity tradeoff.
    \item Most are prone to defects in the crystalline structure.
  \end{itemize}

@@

  Preliminary evidence has shown that cross-linked lyotropic liquid crystal
  (LLC) membranes can be produced at moderate scale and may be capable of 
  performing highly selective separations. 
  \begin{itemize}  
    \item LLCs are amphiphilic molecules that have the ability to self-assemble
    into porous nanostructures \cite{smith_ordered_1997} and can be cross-linked
    to create mechanically strong membrane films with periodic pores on the
    order of 1 nm in diameter \cite{zhou_supported_2005}. 
    \item LLC membrane pores are uniform in size because they are formed by 
    self-assembly. 
    \item Since LLC polymer membranes lack an appreciable pore size distribution,
    they inherently exhibit high selectivity due to their strict molecular weight
    cut-off (MWCO)~\cite{zhou_supported_2005}. 
    \item Additionally, LLC monomers can be salts, and therefore lead to Donnan
    exclusion of ions in solution.\cite{donnan_theory_1995}
%    The membrane gains a net surface charge when counterions from
%  the head groups that line the pore walls escape into the feed solution in an
%  effort to balance the gradients of concentration and electric potential
%  \cite{donnan_theory_1995}.
  \end{itemize}

@@

  The H\textsubscript{II}-phase pore geometry (Figure~\ref{fig:assembly}) has a
  higher theoretical capacity for transport than the Q\textsubscript{I} phase.
  \begin{itemize}
	  \item The H\textsubscript{II} phase forms at room temperature in the 
	  presence of ca.~10 wt\% water and consists of hexagonally packed, 
	  hydrophilic pore columns\cite{smith_ordered_1997}. 
	  % BJC: should rethink the following since we think that thermotropic doesn't exist
	  \item In the absence of water, neat monomer will form the same hexagonal
	  columnar structure which, in the literature, has been referred to as the
	  Col\textsubscript{h} thermotropic phase\cite{feng_scalable_2014}.
  \end{itemize}

@@

  Recently, researchers have learned how to macroscopically align the
  hexagonal domains which has revived research into H\textsubscript{II}-phase LLC
  polymer membranes. 
  \begin{itemize}
    \item In 2014, Feng et al.~showed that one can align Col\textsubscript{h}
    domains, created by the ``dry" monomer Na-GA3C11, using a magnetic field
	with subsequent cross-linking to lock the structure in place\cite{feng_scalable_2014}.
	\item In 2016, Feng et al.~showed that one could also obtain the same result
	by confining the neat monomer between PDMS or glass substrates since hexagonal
	mesophases preferentially anchor perpendicular to both surfaces\cite{feng_thin_2016}.
  \end{itemize}

  Unfortunately, reproducing the work of Feng et al. with the H\textsubscript{II}
  has been an experimental challenge. Therefore, the primary focus of experimental
  research efforts has been with the Q\textsubscript{I} phase.

@@

  A molecular-level understanding of structure of and transport in LLC polymer 
  membranes, enabled by molecular dynamics (MD) simulations, can provide 
  guidelines to reduce the large chemical space available to design
  monomers for creation of separation-specific membranes. 
  \begin{itemize}
    \item Using a sufficiently accurate molecular model, we can directly 
    observe transport of solutes within LLC membrane nanopores and infer
    mechanisms. 
    \item Based on this information we can intelligently design new membranes
    by screening new liquid crystal monomer designs.
    \item Our most promising designs can be tested experimentally.
  \end{itemize}

  \noindent There are five primary objectives of our work.
  \begin{enumerate}

@@

    Useful molecular-level modeling should incorporate a detailed picture 
    of the nanoscopic pore structure, which is crucial to understanding
    the role of monomer structure in solute transport and membrane design.
    We will generate simulated X-ray diffraction patterns from MD trajectories
    and compare them to an experimental 2D wide angle X-ray scattering (WAXS)
    spectrum of a Col\textsubscript{h} phase membrane. We will assess the extent
    to which we can apply our understanding to the H\textsubscript{II} phase,
    as well as systems built with alternate monomers.

@@

    \item Enable easy continuation of our work with a dedicated and 
    well-documented python package.

	Although molecular simulations have become popular
	for studying systems at the atomic level, LLCs used in this context
	have not been heavily investigated. Consequently, much of the 
	analysis developed for this project is not widely applied. Therefore, 
	it is important for us to make available the scripts that reproduce
	the exact results presented in our published papers along with detailed
	documentation of the scripts. This will ensure near-seamless continuation
	of this project and accelerate the development of LLC membranes.

@@

  \textbf{\large Objective 1:} {\large Build and understand the nanoscopic structure of LLC membranes} (\textcolor{green!40!olive}{\textbf{Complete}})

  \noindent Our first task was to develop a procedure for building and equilibrating 
  an atomistic LLC membrane system (see Figure~\ref{fig:assembly}).
  \begin{itemize}
    \item We built a monoclinic unit cell consisting of four pores.
    \item Each pore is composed of monomer columns, where each column
    consists of 20 monomers stacked on top of each other so that the 
    phenyl groups are coplanar with each other and the $xy$ plane. 
    \item The columns are oriented so that the hydrophilic monomer head
    groups face towards the pore center.
    \item We equilibrate the unit cell using a series of simulations
    with head groups held in place by position restraints. 
    \item We gradually reduce the force constant of the position 
    restraints until the system is completely unrestrained.
    \item We carried out 400 ns equilibration simulations.
  \end{itemize}

  \begin{figure}[!htb]
      \centering
      \begin{subfigure}{.3\textwidth}
              \centering
              \includegraphics[width=\textwidth]{NaGA3C11.pdf}
              \caption{}~\label{fig:monomer}
      \end{subfigure}
      \begin{subfigure}{.3\textwidth}
              \centering
              \includegraphics[width=\textwidth]{monomer_twocolor.pdf}
              \caption{}~\label{fig:atomistic_monomer}
      \end{subfigure}
      \begin{subfigure}{0.3\linewidth}
              \centering
              \includegraphics[width=0.6\textwidth]{wedge_thick.pdf}
              \caption{}~\label{fig:wedge}
      \end{subfigure}
              \begin{subfigure}{0.4\linewidth}
              \centering
              \includegraphics[width=0.6\textwidth]{columns.pdf}
              \caption{}~\label{fig:wedge_layer}
      \end{subfigure}
      \begin{subfigure}{0.4\linewidth}
              \centering
              \includegraphics[width=0.6\textwidth]{hexagonal_packing.pdf}
              \caption{}~\label{fig:hex_packing_simple}
      \end{subfigure}
      \caption{(a) The LLC monomer Na-GA3C11 (b) rendered atomistically (c)
      exhibits wedge-like character. (d) Monomers stack on top of each other to
      create columns with short range order, then assemble into pores with
      hydrophilic head groups (blue) facing towards the pore center. (e) The
      pores assemble into hexagonally packed columnar mesophases.}~\label{fig:assembly}
  \end{figure}

  \noindent We tested a number of different initial column architectures. 
  \begin{itemize}
      \item We stacked monomers in parallel displaced and sandwiched 
      configurations, two possible $\pi$-$\pi$ stacking modes.
      \item We also varied the initial distance between stacked monomers, $d$.
      \item We chose values of 3.7 \AA, based on experimental WAXS 
      measurements, as well as 5 \AA as a test of its sensitivity.
  \end{itemize}

  \noindent Our model's geometry is most consistent with experiment for systems
  built with 5 columns per pore and monomers initially stacked 3.7 \AA apart
  (see Figure~\ref{fig:p2p}). 
  \begin{itemize}
    \item We equilibrated systems with 4, 5, 6, 7 and 8 columns per pore.
    \item The pore spacing of 5 column-per-pore systems agree well with experiment.
    \item 6 column-per-pore systems built with $d$ = 5 \AA appear to yield an
    experimentally consistent pore spacing however, the equilibrated distance
    between stacked monomers stay close to 5 \AA which is inconsistent with 
    experiment. 
    \item In general, the equilibrated distance between monomers stays close to
    its initial value. 
  \end{itemize}

@@

  \noindent We further validated the structure of our molecular model by 
  verifying its consistency with five major reflections present in the 
  experimental WAXS pattern.
  \begin{itemize}
    \item The five major reflections and the structural features leading
    to them are summarized in Figure~\ref{fig:WAXS_comparison}.
    \item Using MD trajectories, we simulated X-ray diffraction (XRD) 
    patterns that we could compare to the WAXS data by taking the 
    appropriate cross-section of the time-averaged 3D structure factor. 
    \item None of the models simulated to this point could reproduce R-double.   
  \end{itemize}

  \noindent It is necessary to add a small amount of water to our model in
  order to fully reproduce all features of the WAXS pattern.
  \begin{itemize}
    \item We built our most experimentally consistent structure in the 
    parallel displaced configuration with 1 wt \% water added to the 
    pores. 
    \item This is the only configuration that gives rise to the 
    R-double feature. 
    \item R-double appears because vertically adjacent monomer head groups
    hydrogen bond with shared water molecules, causing them to be drawn 
    closer together.
    \item When multiple pairing interactions occur in series along the 
    same pore axis, the center of masses of the pairs are spaced apart at 
    twice the $\pi$-stacking distance. 
  \end{itemize}

@@

  %BJC: I think this is important to say somewhere
  \noindent On the timescales that we can reasonably simulate, our model
  exhibits slow dynamics. 
  \begin{itemize}
    \item Consequently, there are very few uncorrelated frames in our 
    trajectories which leads to noise in the simulated XRD patterns. 
    \item We can overcome this issue by combining the structure factors
    generated from an ensemble of trajectories produced starting from 
    an ensemble of uncorrelated initial configurations.
  \end{itemize}

@@

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{\textwidth}
  \includegraphics[width=\textwidth]{regional_density_legend.pdf}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{head_group_density.pdf}
        \caption{Head Groups}
        \label{fig:head_groups_regional_density}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{tails_density.pdf}
        \caption{Tails}
        \label{fig:tails_regional_density}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{sodium_density.pdf}
        \caption{Sodium Ions}
        \label{fig:sodium_regional_density}
  \end{subfigure}
  \caption{In all cases, the component radial distribution functions are similar.
      They exhibit a composition gradient transitioning from the hydrophilic to the hydrophobic
          regions. The biggest differences are at r=0 where noise is higher due to
          decreased sampling. The center of the pore is not hollow, but contains sodium ions and
          head groups, even when the system is solvated. This architecture may impede transport in
          the real system in a chemically-dependent manner.
          The solvated system has a lower density of head groups near the
          pore center which is likely due to the swelling that is necessary in order to fit water
          molecules in the pore region.}~\label{fig:overlaid_densities}
  \end{figure}

@@

  \noindent We added additional water to our best Col\textsubscript{h} 
  phase model in order to create the H\textsubscript{II} phase.
  \begin{itemize}
    \item There have been a range of reported water contents used to 
    form the H\textsubscript{II} phase with this particular monomer.
    \item Although Resel et al. commented that the system is likely
    already fully hydrated at 7 wt \% water with additional water
    trapped in defects between mesophases.
    \item Therefore, we built two systems with 5 and 10 wt \% water.
  \end{itemize}

  \noindent We observed that water partitions into the distal tail region.
  \begin{itemize}
    \item We define the distal tail region to be any radial distance 
    greater than 1.5 nm from any pore center. 
    \item There is an approximately 3:2 ratio of water in the pores to
    water in the distal tails. 
    \item Due to the wedge shape of the monomers, the distal tail region
    has a relatively low density leaving space for water molecules to 
    fill.  
  \end{itemize}

@@

  \noindent Water and sodium transport is significantly faster in the less
  crowded 10 wt \% water pores. 
  \begin{itemize}
    \item The MSD of water is about 51 times higher and the MSD of sodium is
    about 49 times higher compared to 5 wt \% system.
    \item In general we observe similar solute transport mechanisms in the 5
    and 10 wt \% water systems but at different time scales.
    \item Therefore, we will focus on the 10 wt \% water system since we
    can collect a larger quantity of mechanistic data from the solute 
    trajectories.
  \end{itemize}

  \noindent We observed transport of 20 different small polar solutes in the
  pores of our model. 
  \begin{itemize}
    \item We created a separate initial configuration for each solute studied.
    \item We placed 6 solutes, equally spaced in $z$, in each nanopore.
    \item We allowed the solutes to simulate for 1 microsecond. 
  \end{itemize}

  \noindent The MSDs are not a monotonic function of solute size.
  \begin{itemize}
    \item In Figure~\ref{fig:all_msds_10wt}, we plotted the average
    MSD of each solute. 
    \item In Figure~\ref{fig:msd_radius_10wt}, we plotted solute
    MSDs against their molecular radius.
    \item We also plotted theoretical curves which illustrate the 
    expected MSD of each solute if they were to travel unhindered.
    \item We ensured that the Stokes-Einstein equation with the
    correction factor of Gierer and Wirtz passed through methanol 
    in order to give an approximate frame of reference. 
    \item Methanol is small and therefore can travel relatively
    unhindered compared to all other solutes.
    \item The theoretical line can be used as an approximate boundary
    between subdiffusive, Brownian and superdiffusive behavior.
  \end{itemize}

>>>>

  % re-word. This paragraph is verbatim from structure paper
  More highly selective nanoporous membranes would be useful for 
  performing complex aqueous separations with seawater and various
  types of wastewater.
  \begin{itemize}
    \item For example, sodium chloride and boron in seawater 
    \cite{fritzmann_state---art_2007} and organic micropollutants found in
    municipal and industrial wastewaters \cite{schwarzenbach_challenge_2006}
    represent just a few of the diverse contaminants of water sources. 
    \item By efficiently separating contaminants from feed solutions with
    highly selective membranes, it is possible to reduce the number of 
    required membrane passes and post-treatment steps needed for a given 
    filtration process \cite{werber_materials_2016}, thus lowering cost
    and energy requirements. 
    \item Additionally, one can extract valuable resources from the 
    feed streams. For example, flowback water produced during hydraulic
    fracturing of shale formations contains dissolved species such as acetate
    whose extraction has economic value \cite{dischinger_application_2017}.
  \end{itemize}

  Reverse osmosis (RO) and nanofiltration (NF) are two prevailing membrane
  filtration processes that can be used to separate solutes on the order of
  1 nm in size and smaller, including ions.
  \begin{itemize}  
    \item Both apply hydraulic pressure to the feed solution in order to 
    overcome osmotic pressure and force water and unfiltered components 
    through the membrane.
    \item RO membranes are typically thin film composite with a porous mechanical 
    support layer and a thin but dense polymer matrix active layer where separations
    occur.\cite{jeong_interfacial_2007}
    \item RO separates solutes based on the solute's ability to dissolve into
    and diffuse through the tortuous pathways available in the membrane's dense 
    active layer.
    \item RO offers high selectivity at the cost of relatively high energy 
    requirements since one must apply hydraulic pressure of up to 100 bar in 
    order to achieve an economical flux~\cite{van_der_bruggen_review_2003}.  % get number
    \item In contrast to RO membranes, NF membranes have explicit pores on the
    order of 1 nm in size. 
    \item Typically, separations are achieved based on size exclusion and
    charge exclusion if the membrane's surface has a net charge.
    \item NF membranes require significantly less applied pressure in order
    to achieve solute flux comparable to RO.
    \item Unfortunately, conventional synthesis processes, such as 
    phase-inversion\cite{smolders_microstructures_1992} are stochastic in
    nature which yields pores that are polydisperse in size.\cite{werber_materials_2016} % name some synthesis processes
    \item Pore size polydispersity is detrimental to membrane selectivity
  \end{itemize}

@@

  Nanostructured membranes attempt to overcome the permeability-selectivity 
  tradeoff through intelligent design at the molecular level.
  \begin{itemize}
	\item Graphene sheets, carbon nanotubes (CNTs) and zeolites are three
	highly studied nanostructured technologies.
  \end{itemize}

  Ultrathin-film graphene and graphene oxide membranes are a very active
  area of research because they offer potential for extremely high 
  permeability membranes.
  \begin{itemize}
    \item An ideal graphene membrane is a 2D material consisting of a
    single layer of graphene.\cite{humplik_nanostructured_2011}
    \item Synthesis of a single layer of graphene at a large scale without
    introducing defects is a challenge that has not yet been overcome.
    \item Multilayered graphene membranes are a potentially easier and
    more scalable technology which has garnered increasing interest in
    recent years.\cite{cohen-tanugi_multilayer_2016,wei_multilayred_2018}
  \end{itemize}

  Carbon nanotubes (CNTs) have shown promise as aqueous separations membranes due
  to unprecedentedly fast water transport.\cite{humplik_nanostructured_2011,hummer_water_2001} 
  \begin{itemize}
    \item Practically, dispersing CNTs into a polymer matrix is extremely difficult
    because they tend to agglomerate due to Van der Waals forces.
    \item Functionalization of the CNT walls has been heavily investigated 
    as a way to overcome this issue. 
    \item Alignment of the CNTs into an array feasible as a nanoporous
    membrane presents a further challenge to CNT membranes.\cite{sahoo_polymer_2010}
  \end{itemize}

  Zeolite-coated ceramic membranes offer the potential for permeabilities
  comparable to ultrafiltration with selectivities as good as NF and RO. 
  \begin{itemize}
    \item Zeolites have highly uniform nm-sized crystalline structures with
    cage-like cavities that allow movement and trapping of small solutes. 
    \item The crystalline frameworks are typically formed by networks of silicon
    and aluminum each attached to 4 oxygen atoms in a tetrahedral arrangement. 
    \item One can replace the silicon and aluminum atoms via ion exchange 
    in order to control the size of the cavities and hence its molecular-seiving properties.
    \item A number of studies have tested the permeability and sodium salt
    rejection of various zeolite membranes, however none have fully
    overcome the permeability-selectivity tradeoff.
    \item Most are prone to defects in the crystalline structure.\cite{pendergast_review_2011}
  \end{itemize}

@@

  Preliminary evidence has shown that cross-linked lyotropic liquid crystal
  (LLC) membranes can be produced at moderate scale and may be capable of 
  performing highly selective separations. 
  \begin{itemize}  
    \item LLCs are amphiphilic molecules that have the ability to self-assemble
    into porous nanostructures\cite{smith_ordered_1997} that
    can be cross-linked to create mechanically strong membrane films with 
    periodic uniform-sized pores on the order of 1 nm in diameter \cite{zhou_supported_2005}. 
    \item Since LLC polymer membranes lack an appreciable pore size distribution,
    they inherently exhibit high selectivity due to their strict molecular weight
    cut-off (MWCO)~\cite{zhou_supported_2005}. 
    \item Additionally, LLC monomers can be salts, and therefore lead to Donnan
    exclusion of ions in solution.\cite{donnan_theory_1995}
%    The membrane gains a net surface charge when counterions from
%  the head groups that line the pore walls escape into the feed solution in an
%  effort to balance the gradients of concentration and electric potential
%  \cite{donnan_theory_1995}.
  \end{itemize}

@@

@@

%  The H\textsubscript{II}-phase pore geometry (Figure~\ref{fig:assembly}) has a
%  high theoretical capacity for transport than the Q\textsubscript{I} phase.
%  \begin{itemize}
%	  \item The H\textsubscript{II} phase forms at room temperature in the 
%	  presence of ca.~10 wt\% water and consists of hexagonally packed, 
%	  hydrophilic pore columns\cite{smith_ordered_1997}. 
%%	  \item The hexagonal, straight-pore geometry provides a dense array
%%	  of hydrophilic channels and short pathways for solutes to travel.
%	  % BJC: should rethink the following since we think that thermotropic doesn't exist
%	  \item In the absence of water, neat monomer will form the same hexagonal
%	  columnar structure which, in the literature, has been referred to as the
%	  Col\textsubscript{h} thermotropic phase\cite{feng_scalable_2014}.
%  \end{itemize}

  Recently, researchers have learned how to macroscopically align the
  hexagonal domains which has revived research into H\textsubscript{II}-phase LLC
  polymer membranes. 
  \begin{itemize}
    \item In 2014, Feng et al.~showed that one can align Col\textsubscript{h}
    domains, a temperature-dependent hexagonal phase created by neat LLC monomers,
    using a magnetic field with subsequent cross-linking to lock the structure 
    in place\cite{feng_scalable_2014}.
	\item In 2016, Feng et al.~showed that one could also obtain the same result
	by confining the neat monomer between PDMS or glass substrates since hexagonal
	mesophases preferentially anchor perpendicular to both surfaces\cite{feng_thin_2016}.
  \end{itemize}

  Unfortunately, reproducing the work of Feng et al. with the hydrated 
  H\textsubscript{II} phase has been an experimental challenge. Therefore, the
  primary focus of experimental research efforts has been with the Q\textsubscript{I} phase.

@@

  A molecular-level understanding of structure and transport in LLC polymer 
  membranes, enabled by molecular dynamics (MD) simulations, can provide 
  guidelines to reduce the large chemical space available to design
  monomers for creation of separation-specific membranes. 
  \begin{itemize}
    \item Using a sufficiently accurate molecular model, we can directly 
    observe transport of solutes within LLC membrane nanopores and infer
    mechanisms. 
    \item Based on this information we can intelligently design new membranes
    by screening new liquid crystal monomer designs.
    \item Our most promising designs can be tested experimentally.
  \end{itemize}

  \noindent There are four primary objectives of this PhD research.
  \begin{enumerate}

@@

    Useful molecular-level modeling should incorporate a detailed picture 
    of the nanoscopic pore structure, which is crucial to understanding
    the role of monomer structure in solute transport and membrane design.
    We will generate simulated X-ray diffraction (XRD) patterns from MD trajectories
    and compare them to an experimental 2D wide angle X-ray scattering (WAXS)
    spectrum of a Col\textsubscript{h} phase membrane. We will assess the extent
    to which we can apply our understanding to the H\textsubscript{II} phase,
    as well as systems built with alternate monomers.

@@

%    \item Enable easy continuation of our work with a dedicated and 
%    well-documented python package.
%    
%	Although molecular simulations have become popular
%	for studying systems at the atomic level, LLCs used in this context
%	have not been heavily investigated. Consequently, much of the 
%	analysis developed for this project is not widely applied. Therefore, 
%	it is important for us to make available the scripts that reproduce
%	the exact results presented in our published papers along with detailed
%	documentation of the scripts. This will ensure near-seamless continuation
%	of this project and accelerate the development of LLC membranes.

@@

  \textbf{\large Objective 1:} {\large Build and understand the nanoscopic structure
  of the H\textsubscript{II} phase} (\textcolor{green!40!olive}{\textbf{Complete}})

  \noindent Our first task was to develop a procedure for building and equilibrating 
  an atomistic LLC membrane system (see Figure~\ref{fig:assembly}).
  \begin{itemize}
    \item We built a monoclinic unit cell consisting of four pores.
    \item Each pore is composed of monomer columns, where each column
    consists of 20 monomers stacked on top of each other so that the 
    phenyl groups are coplanar with each other and the $xy$ plane. 
    \item The columns are oriented so that the hydrophilic monomer head
    groups face towards the pore center.
    \item We did not add any water to the initial configuration because we
    compared our structure to experimental data for a system claimed to be
    synthesized dry.
    \item We equilibrated the unit cell using a series of simulations
    with head groups held in place by position restraints. 
    \item We gradually reduce the force constant of the position 
    restraints until the system is completely unrestrained.
    \item We carried out 400 ns equilibration simulations.

  \end{itemize}

  \begin{figure}[!htb]
      \centering
      \begin{subfigure}{.325\textwidth}
              \centering
              \vspace{0.5cm}
              \includegraphics[width=\textwidth]{NaGA3C11.pdf}  %BJC: can color code this once I figure out how
              \vspace{0.05cm}
              \caption{}~\label{fig:monomer}
      \end{subfigure}
%      \begin{subfigure}{.3\textwidth}
%              \centering
%              \includegraphics[width=\textwidth]{monomer_twocolor.pdf}
%              \caption{}~\label{fig:atomistic_monomer}
%      \end{subfigure}
%      \begin{subfigure}{0.3\linewidth}
%              \centering
%              \includegraphics[width=0.6\textwidth]{wedge_thick.pdf}
%              \caption{}~\label{fig:wedge}
%      \end{subfigure}
              \begin{subfigure}{0.325\linewidth}
              \centering
              \includegraphics[width=0.6\textwidth]{columns.pdf}
              \caption{}~\label{fig:wedge_layer}
      \end{subfigure}
      \begin{subfigure}{0.325\linewidth}
              \centering
              \includegraphics[width=0.6\textwidth]{hexagonal_packing.pdf}
              \caption{}~\label{fig:hex_packing_simple}
      \end{subfigure}
      \caption{(a) The LLC monomer Na-GA3C11 exhibits wedge-like character. (b) 
      Monomers stack on top of each other to create columns with short range order,
      then assemble into pores with hydrophilic head groups (blue) facing towards 
      the pore center. (c) The pores assemble into hexagonally packed columnar
      mesophases.}~\label{fig:assembly}
  \end{figure}

  \noindent We tested a number of different initial column architectures. 
  \begin{itemize}
      \item We stacked monomers in parallel displaced and sandwiched 
      configurations, two possible $\pi$-$\pi$ stacking modes.\cite{sinnokrot_estimates_2002}
      \item We also varied the initial distance between stacked monomers, $d$.
      \item We chose values of 3.7 \AA, based on experimental WAXS 
      measurements, as well as 5 \AA as a test of its sensitivity.
  \end{itemize}

  \noindent Our model's geometry is most consistent with experiment for systems
  built with 5 columns per pore and monomers initially stacked 3.7 \AA~apart
  (see Figure~\ref{fig:p2p}). 
  \begin{itemize}
    \item We equilibrated systems with 4, 5, 6, 7 and 8 columns per pore.
    \item The pore spacing of 5 column-per-pore systems agree well with experiment.
    \item 6 column-per-pore systems built with $d$ = 5 \AA~appear to yield an
    experimentally consistent pore spacing however, the equilibrated distance
    between stacked monomers stay close to 5 \AA~which is inconsistent with 
    experiment. 
    \item In general, the equilibrated distance between monomers stays close to
    its initial value. 
  \end{itemize}

@@

  \noindent We further validated the structure of our molecular model by 
  verifying its consistency with five major reflections present in the 
  experimental WAXS pattern.
  \begin{itemize}
    \item The five major reflections and the structural features leading
    to them are summarized and described in the caption of Figure~\ref{fig:WAXS_comparison}.
    \item Using MD trajectories, we simulated X-ray diffraction (XRD) 
    patterns that we could compare to the WAXS data by taking the 
    appropriate cross-section of the time-averaged 3D structure factor. 
    \item None of the models simulated to this point could reproduce R-double.   
  \end{itemize}

  \noindent It is necessary to add a small amount of water to our model in
  order to fully reproduce all features of the WAXS pattern (see 
  Figure~\ref{fig:solvated_XRD}).
  \begin{itemize}
    \item We built our most experimentally consistent structure in the 
    parallel displaced configuration with 1 wt\% water added to the 
    pores. 
    \item This is the only configuration that gives rise to the 
    R-double feature. 
    \item R-double appears because vertically adjacent monomer head groups
    hydrogen bond with shared water molecules, causing them to be drawn 
    closer together.
    \item When multiple pairing interactions occur in series along the 
    same pore axis, the center of masses of the pairs are spaced apart at 
    twice the $\pi$-stacking distance. 
    \item The necessity of water in our model suggests that the membrane
    synthesized by Feng et al. was slightly hydrated due to water molecules
    leached from surroundings by the hydroscopic monomers. 
  \end{itemize}

@@

  %BJC: I think this is important to say somewhere
  %BJC: but I don't have an ensemble of 1 wt % systems to fully prove that with.
  %BJC: I could show cross-sections of R-pi, but that would require more context. 
  \noindent On the timescales that we can reasonably simulate, our model
  exhibits slow dynamics. 
  \begin{itemize}
    \item Consequently, there are very few uncorrelated frames in our 
    trajectories which leads to noise in the simulated XRD patterns. 
    \item We can overcome this issue by combining the structure factors
    generated from an ensemble of trajectories produced starting from 
    an ensemble of uncorrelated initial configurations.
  \end{itemize}

@@

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{\textwidth}
  %BJC: probably change this legend to `d = x Angstroms` instead of 
  % disordered/ordered, since I don't have space to talk about ordered / disorderd configurations (requires 
  % nematic order parameter to give a complete justification of the names)
  \includegraphics[width=\textwidth]{regional_density_legend.pdf}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{head_group_density.pdf}
        \caption{Head Groups}
        \label{fig:head_groups_regional_density}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{tails_density.pdf}
        \caption{Tails}
        \label{fig:tails_regional_density}
  \end{subfigure}
  \begin{subfigure}{0.32\textwidth}
        \includegraphics[width=1\linewidth]{sodium_density.pdf}
        \caption{Sodium Ions}
        \label{fig:sodium_regional_density}
  \end{subfigure}
  \caption{In all cases, the component radial distribution functions are similar.
      They exhibit a composition gradient transitioning from the hydrophilic to the hydrophobic
          regions. The biggest differences are at r=0 where noise is higher due to
          decreased sampling. The center of the pore is not hollow, but contains sodium ions and
          head groups, even when the system is solvated. This architecture may impede transport in
          the real system in a chemically-dependent manner.
          The solvated system has a lower density of head groups near the
          pore center which is likely due to the swelling that is necessary in order to fit water
          molecules in the pore region.}~\label{fig:overlaid_densities}
  \end{figure}

@@

  \noindent We added additional water to our most experimentally consistent
  structural model in order to create a higher water content H\textsubscript{II}
  phase model.
  \begin{itemize}
    \item There have been a range of reported water contents used to 
    form the H\textsubscript{II} phase with this particular monomer.
    \item Although Resel et al. commented that the system is likely
    already fully hydrated at 7 wt \% water with additional water
    trapped in defects between mesophases.\cite{resel_structural_2000}
    \item Therefore, we built two systems with 5 and 10 wt\% water.
  \end{itemize}

  \noindent We observed that water partitions into the distal tail region.
  \begin{itemize}
    \item We define the distal tail region to be any radial distance 
    greater than 1.5 nm from any pore center. 
    \item There is approximately a 3:2 ratio of water in the pores to
    water in the distal tails. 
    \item Due to the wedge shape of the monomers, the distal tail region
    has a relatively low density leaving space for water molecules to 
    fill.  
  \end{itemize}

@@

  \noindent Water and sodium transport is significantly faster in the less
  crowded 10 wt\% water pores. 
  \begin{itemize}
    \item The mean squared displacement (MSD) of water is about 51 times 
    higher and the MSD of sodium is about 49 times higher compared to the 5 wt\% system.
    \item In general we observe similar solute transport mechanisms in the 5
    and 10 wt\% water systems but on different time scales.
    %BJC: might as well leave 5 wt% data in there.
%    \item Therefore, we focus the remainder of this discussion on the 10 wt\%
%    water system since we were able to collect a larger quantity of mechanistic
%    data from the solute trajectories.
  \end{itemize}

  % BJC: do you think I can get away with an appendix with the solute table + abbreviations that
  % doesn't count towards the page limit?
  \noindent We observed transport of 20 different small polar solutes in the
  pores of our model. 
  \begin{itemize}
    \item We created a separate initial configuration for each solute studied and
    placed 6 solutes, equally spaced in $z$, in each nanopore.
    \item After a brief equilibration, we allowed the solutes to simulate for
    1 microsecond. 
  \end{itemize}
